{\bf Affine.m} -- {\it Mathematica} package for computations in representation theory of finite-dimensional and affine Lie algebras

### 2011-06-16 doc/paper.tex
Minor changes

Each $\gf$-module is also an $\af$-module, although
$L^{\mu}_{\gf}$ is not irreducible as $\af$-module. It can be
decomposed into the sum of irreducible $\af$-modules:
\begin{equation}
  \label{eq:16}
  L^{\mu}_{\gf}=\bigoplus_{\nu}b^{\mu}_{\nu}L^{\nu}_{\af}
\end{equation}
Coefficients in this decomposition are called the branching
coefficients. Using the Weyl character formula we can write a
recurrent relation \cite{2010arXiv1007.0318L} for branching
coefficients $k_{\xi }^{\left( \mu \right) }$ corresponding to an
injection $\af\hookrightarrow \gf$:
\begin{equation}
\begin{array}{c}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{u\in U}\epsilon (u)\;\dim \left( L_{\afb}^{\mu _{\af%
_{\perp }}\left( u\right) }\right) \delta _{\xi -\gamma _{0},\pi _{%
\widetilde{\af}}(u(\mu +\rho )-\rho )}+\right.  \\
\left. +\sum_{\gamma \in \Gamma _{\widetilde{\af}\rightarrow \gf%
}}s\left( \gamma +\gamma _{0}\right) k_{\xi +\gamma }^{\left( \mu \right)
}\right) .
\end{array}
\label{recurrent-rel}
\end{equation}
The recursion is goverened by the set $\Gamma _{\widetilde{\af}\rightarrow \gf}$ called the injection fan. The latter is defined by the
carrier set $\left\{ \xi \right\} _{\af\rightarrow \gf}$ for the
coefficient function $s(\xi )$
\begin{equation*}
\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}:=\left\{
\xi \in P_{\widetilde{\af}}|s(\xi )\neq 0\right\}
\end{equation*}
appearing in the expansion
\begin{equation}
\prod_{\alpha \in \Delta ^{+}\setminus \Delta _{\bot }^{+}}\left( 1-e^{-\pi
_{\widetilde{\af}}\alpha }\right) ^{\mathrm{mult}(\alpha )-\mathrm{mult}%
_{\af}(\pi _{\widetilde{\af}}\alpha )}=-\sum_{\gamma \in P_{%
\widetilde{\af}}}s(\gamma )e^{-\gamma };\quad
\end{equation}
The weights in $\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \frak{%
g}}$ are to be shifted by $\gamma _{0}$ -- the lowest vector in $\left\{ \xi
\right\} $ -- and the zero element is to be eliminated:
\begin{equation}
\Gamma _{\widetilde{\af}\rightarrow \gf}=\left\{ \xi -\gamma
_{0}|\xi \in \left\{ \xi \right\} \right\} \setminus \left\{ 0\right\} .
\end{equation}
The formula (\ref{eq:14}) is a particular case of recurrent relation for branching coefficients (\ref{recurrent-rel}) in the case of Cartan subalgebra $\af=\hfg$.

@@

\section{Computational algorithms}
\label{sec:comp-algor}

@@

\item  Construct the set $\Gamma _{\af\rightarrow \frak{g}}$. Relation
 (\ref{eq:6}) defines the sign function
 $s(\gamma)$ and the set $\Phi_{\af\subset \frak{g}}$ where the lowest weight
 $\gamma_0$ is to be subtracted to get the fan (\ref{fan-defined}):
 $\Gamma _{\af\rightarrow \frak{g}}=\left\{ \xi -\gamma _{0}|\xi \in \Phi _{%
\af\subset \frak{g}}\right\} \setminus \left\{ 0\right\}$.

@@

\item  Calculate the anomalous branching coefficients using the
recurrent relation (\ref{recurrent-relation}) and select among them those
corresponding to the weights in the main Weyl
chamber $\overline{C_{\af}}$.
\end{enumerate}

@@

\begin{figure}[h]
  \label{freudenthal-racah-times}
  \noindent\centering{
 %   \includegraphics[width=80mm]{figure1.eps}
  }
  \caption{Freudenthal run time with the number of weights in $\bar C$.}

@@

For fundamental representations of simple Lie algebras it is sometimes possible to get analytic result for the dependence of decomposition coefficients on $N$ (See \cite{2011LyakhovskyPostnova}). Our code give numerical values and can be used to check this analytic results.

>>>>

Each $\gf$-module is also an $\af$-module, although
$L^{\mu}_{\gf}$ is not irreducible as $\af$-module. It can be
decomposed into the sum of irreducible $\af$-modules:
\begin{equation}
  \label{eq:16}
  L^{\mu}_{\gf}=\bigoplus_{\nu}b^{\mu}_{\nu}L^{\nu}_{\af}
\end{equation}
Coefficients in this decomposition are called the branching
coefficients. Using the Weyl character formula we can write a
recurrent relation \cite{2010arXiv1007.0318L} for branching
coefficients $k_{\xi }^{\left( \mu \right) }$ corresponding to an
injection $\af\hookrightarrow \gf$:
\begin{equation}
\begin{array}{c}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{u\in U}\epsilon (u)\;\dim \left( L_{\afb}^{\mu _{\af%
_{\perp }}\left( u\right) }\right) \delta _{\xi -\gamma _{0},\pi _{%
\widetilde{\af}}(u(\mu +\rho )-\rho )}+\right.  \\
\left. +\sum_{\gamma \in \Gamma _{\widetilde{\af}\rightarrow \gf%
}}s\left( \gamma +\gamma _{0}\right) k_{\xi +\gamma }^{\left( \mu \right)
}\right) .
\end{array}
\label{recurrent-rel}
\end{equation}
The recursion is goverened by the set $\Gamma _{\widetilde{\af}\rightarrow \gf}$ called the injection fan. The latter is defined by the
carrier set $\left\{ \xi \right\} _{\af\rightarrow \gf}$ for the
coefficient function $s(\xi )$
\begin{equation*}
\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}:=\left\{
\xi \in P_{\widetilde{\af}}|s(\xi )\neq 0\right\}
\end{equation*}
appearing in the expansion
\begin{equation}
\prod_{\alpha \in \Delta ^{+}\setminus \Delta _{\bot }^{+}}\left( 1-e^{-\pi
_{\widetilde{\af}}\alpha }\right) ^{\mathrm{mult}(\alpha )-\mathrm{mult}%
_{\af}(\pi _{\widetilde{\af}}\alpha )}=-\sum_{\gamma \in P_{%
\widetilde{\af}}}s(\gamma )e^{-\gamma };\quad
\end{equation}
The weights in $\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}$ are to be shifted by $\gamma _{0}$ -- the lowest vector in $\left\{ \xi
\right\} $ -- and the zero element is to be eliminated:
\begin{equation}
\Gamma _{\widetilde{\af}\rightarrow \gf}=\left\{ \xi -\gamma
_{0}|\xi \in \left\{ \xi \right\} \right\} \setminus \left\{ 0\right\} .
\end{equation}
The formula (\ref{eq:14}) is a particular case of recurrent relation for branching coefficients (\ref{recurrent-rel}) in the case of Cartan subalgebra $\af=\hfg$.

@@

\subsection{Modules}
\label{sec:modules}

{\bf Affine.m} can be used to study different kinds of modules, e.g. Verma modules, irreducible modules.  We need datastructure to represent generic module of Lie algebra $\gf$. This datastructure is \lstinline{module}. Module properties can be deduced from its set of singular weights using Weyl character formulae \eqref{eq:11},\eqref{eq:12},\eqref{eq:18},\eqref{eq:13}. Set of singular weights can have Weyl symmetry. It can be symmetry with respect to Weyl group $W_{\gf}$ or with respect to some subalgebra $W_{\af}$ as in the case of parabolic Verma modules. Then it is possible to study only main Weyl chamber $C_{\af}$. So generic constructor for \lstinline{module} datastructure is \lstinline{makeModule[rs_?rootSystemQ][singWeights_formalElement,subs_?rootSystemQ|emptyRootSystem[],opts___?OptionQ]}. Here \lstinline{rs} is root system of Lie algebra $\gf$,\lstinline{singWeights} is the set of singular weights,  \lstinline{subs} is root system corresponding to Weyl group $W_{\af}$ which is the symmetry of the set of singular weights.
There are several constructors for different highest weight modules:
\begin{lstlisting}[mathescape=true]
vm=makeVermaModule[$B_{2}$][weight[$B_{2}$][2,1]];
pm=makeParabolicVermaModule[$B_{2}$][weight[$B_{2}$][2,1],{1}];
im=makeIrreducibleModule[$B_{2}$][weight[$B_{2}$][2,1]];
drawPlaneProjection[2,1,{character[vm],character[pm],character[im]} ]
\end{lstlisting}
\section{Computational algorithms}
\label{sec:comp-algor}

@@

\item  Construct the set $\Gamma _{\af\rightarrow \frak{g}}$. Relation
 (\ref{eq:6}) defines the sign function
 $s(\gamma)$ and the set $\Phi_{\af\subset \frak{g}}$ where the lowest weight
 $\gamma_0$ is to be subtracted to get the fan:
 $\Gamma _{\af\rightarrow \frak{g}}=\left\{ \xi -\gamma _{0}|\xi \in \Phi _{%
\af\subset \frak{g}}\right\} \setminus \left\{ 0\right\}$.

@@

\item  Calculate the anomalous branching coefficients using the
recurrent relation (\ref{recurrent-rel}) and select among them those
corresponding to the weights in the main Weyl
chamber $\overline{C_{\af}}$.
\end{enumerate}

@@

\begin{figure}[h]
%  \label{freudenthal-racah-times}
  \noindent\centering{
 %   \includegraphics[width=80mm]{figure1.eps}
  }
  \caption{Freudenthal run time with the number of weights in $\bar C$.}

@@

For fundamental representations of simple Lie algebras it is sometimes possible to get analytic result for the dependence of decomposition coefficients on $N$ (See \cite{LyakhovskyPostnova2011}). Our code give numerical values and can be used to check this analytic results.
